### 2025-03-13 paper/quarto_ecml/paper.tex
Review experiments

Our key outcome of interest is how well models perform with respect to
explainability (Definition~\ref{def-explainability}): to this end, we
focus primarily on the plausibility and cost of faithfully generated
counterfactuals at test time. To measure the cost of counterfactuals, we
follow the standard convention of using distances (\(\ell_1\)-norm)
between factuals and counterfactuals as a proxy. For plausibility, we
assess how similar counterfactuals are to observed samples in the target
domain. We rely on the distance-based metric used by
\cite{altmeyer2024faithful},

@@

where \(\mathbf{X}^\prime\) denotes a set of multiple counterfactuals
and \(\text{MMD}(\cdot)\) is an unbiased estimate of the squared
population maximum mean discrepancy \cite{gretton2012kernel}. The metric
in Equation~\ref{eq-impl-div} is equal to zero iff
\(\mathbf{X}^\prime=\mathbf{X}^+\).

@@

We run the experiments with three CE generators: \emph{Generic} of
\cite{wachter2017counterfactual} as a simple baseline approach,
\emph{REVISE} \cite{joshi2019realistic} that aims to generate plausible
counterfactuals using a surrogate Variational Autoencoder (VAE), and
\emph{ECCo}---the generator of \cite{altmeyer2023faithful} but without
the conformal prediction component---as a method that directly targets
both faithfulness and plausibility of the CEs.

@@

We begin the discussion by addressing the direct extensions of the
counterfactual training approach in Section~\ref{sec-approach}. Then, we
look at its broader limitations and challenges in
Section~\ref{sec-limitations}.

>>>>

Our key outcome of interest is how well do models perform with respect
to explainability (Definition~\ref{def-explainability}). To this end, we
focus primarily on the plausibility and cost of faithfully generated
counterfactuals at test time. To measure the cost of counterfactuals, we
follow the standard convention of using distances (\(\ell_1\)-norm)
between factuals and counterfactuals as a proxy. For plausibility, we
assess how similar counterfactuals are to observed samples in the target
domain. We rely on the distance-based metric used by
\cite{altmeyer2024faithful},

@@

where \(\mathbf{X}^\prime\) denotes a set of multiple counterfactuals
and \(\text{MMD}(\cdot)\) is an unbiased estimate of the squared
population maximum mean discrepancy \cite{gretton2012kernel}. The metric
in Equation~\ref{eq-impl-div} is equal to zero iff the two distributions
are the same, \(\mathbf{X}^\prime=\mathbf{X}^+\).

@@

We run the experiments with three gradient-based generators:
\emph{Generic} of \cite{wachter2017counterfactual} as a simple baseline
approach, \emph{REVISE} \cite{joshi2019realistic} that aims to generate
plausible counterfactuals using a surrogate Variational Autoencoder
(VAE), and \emph{ECCo}---the generator of \cite{altmeyer2023faithful}
but without the conformal prediction component---as a method that
directly targets both faithfulness and plausibility of the CEs.

@@

We first address the direct extensions of the counterfactual training
approach in Section~\ref{sec-approach}. Then, we look at its limitations
and challenges in Section~\ref{sec-limitations}.

### 2025-03-13 paper/sections/discussion.qmd
Review experiments

We begin the discussion by addressing the direct extensions of the counterfactual training approach in @sec-approach. Then, we look at its broader limitations and challenges in @sec-limitations.

>>>>

We first address the direct extensions of the counterfactual training approach in @sec-approach. Then, we look at its limitations and challenges in @sec-limitations.

### 2025-03-13 paper/sections/experiments.qmd
Review experiments

Our key outcome of interest is how well models perform with respect to explainability (@def-explainability): to this end, we focus primarily on the plausibility and cost of faithfully generated counterfactuals at test time. To measure the cost of counterfactuals, we follow the standard convention of using distances ($\ell_1$-norm) between factuals and counterfactuals as a proxy. For plausibility, we assess how similar counterfactuals are to observed samples in the target domain. We rely on the distance-based metric used by @altmeyer2024faithful,

@@

where $\mathbf{X}^\prime$ denotes a set of multiple counterfactuals and $\text{MMD}(\cdot)$ is an unbiased estimate of the squared population maximum mean discrepancy [@gretton2012kernel]. The metric in @eq-impl-div is equal to zero iff $\mathbf{X}^\prime=\mathbf{X}^+$.

@@

We run the experiments with three CE generators: *Generic* of @wachter2017counterfactual as a simple baseline approach, *REVISE* [@joshi2019realistic] that aims to generate plausible counterfactuals using a surrogate Variational Autoencoder (VAE), and *ECCo*---the generator of @altmeyer2023faithful but without the conformal prediction component---as a method that directly targets both faithfulness and plausibility of the CEs.

>>>>

Our key outcome of interest is how well do models perform with respect to explainability (@def-explainability). To this end, we focus primarily on the plausibility and cost of faithfully generated counterfactuals at test time. To measure the cost of counterfactuals, we follow the standard convention of using distances ($\ell_1$-norm) between factuals and counterfactuals as a proxy. For plausibility, we assess how similar counterfactuals are to observed samples in the target domain. We rely on the distance-based metric used by @altmeyer2024faithful,

@@

where $\mathbf{X}^\prime$ denotes a set of multiple counterfactuals and $\text{MMD}(\cdot)$ is an unbiased estimate of the squared population maximum mean discrepancy [@gretton2012kernel]. The metric in @eq-impl-div is equal to zero iff the two distributions are the same, $\mathbf{X}^\prime=\mathbf{X}^+$.

@@

We run the experiments with three gradient-based generators: *Generic* of @wachter2017counterfactual as a simple baseline approach, *REVISE* [@joshi2019realistic] that aims to generate plausible counterfactuals using a surrogate Variational Autoencoder (VAE), and *ECCo*---the generator of @altmeyer2023faithful but without the conformal prediction component---as a method that directly targets both faithfulness and plausibility of the CEs.
